---
author: "Rebecca Kee"
date: "8/1/2019"
output: 
    pdf_document: default
    html_document:
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Statistical Inference Course Project Part 2

## Overview
In the second part of this project, we're going to analyze the ToothGrowth data in the R datasets package.

### Instructions 
1. Load the ToothGrowth data and perform some basic exploratory data analyses
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose, using only techniques from class.
4. State your conclusions and the assumptions needed for your conclusions.

## Exploratory Data Analyses

First, load the data and examine its structure
```{r}
# Load required packages
library(ggplot2)

# Load data
data("ToothGrowth")

# Examine structure
str(ToothGrowth)
```

Next, let's do a basic summary of the data
```{r}
summary(ToothGrowth)
```

Let's plot the summary of tooth growth by delivery method and dose
```{r}
ggplot(ToothGrowth, aes(x = dose, y = len)) + 
  geom_boxplot(aes(fill = factor(dose)), show.legend = FALSE) +
  facet_grid(. ~ supp) +
  labs(x = "Dose (mg/day)", 
       y = "Length of odontoblasts cells", 
       title = "Tooth growth by delivery method and dose"
       ) +
  ylim(range(pretty(c(0, ToothGrowth$len)))) + 
  theme_bw() + 
  scale_fill_brewer(palette="Pastel1") +
  theme(plot.title = element_text(hjust = 0.5))

```

## Hypothesis Testing

### Does Vitamin C delivery method affect tooth growth? 

The first hypothesis we will test is as follows: 
$H_0$: The Vitamin C delivery method (orange juice vs ascorbic acid) does not affect tooth growth (i.e. The difference in means between both methods equals 0).
$H_1$: The Vitamin C delivery method (orange juice vs ascorbic acid) affects tooth growth. (i.e. The difference in means between both methods does not equal 0).

Run t-test
```{r}
t.test(len~supp, data = ToothGrowth)
```
Given that the p-value is about 0.06 and the confidence interval (CI) contains 0, we fail to reject the null hypothesis. The delivery method of Vitamin C does not seem to affect tooth growth.

### Does Vitamin C dosing affect tooth growth?

The next hypothesis we will test is as follows: 
$H_0$: The Vitamin C dose does not affect tooth growth (i.e. The difference in means between doses equals 0).
$H_1$: The Vitamin C dose affects tooth growth (i.e. The difference in means between doses does not equal 0).

Because there are 3 different doses, we will run 2 t-tests. 
```{r}
options(scipen = 999)

# Run t-test for doses 0.5 mg/day vs 1.0 mg/day
dose.subset1 <- subset(ToothGrowth, ToothGrowth$dose %in% c(0.5, 1.0))
ttest1 <- t.test(len~dose, data = dose.subset1)
ttest1
```


```{r}
# Run t-test for doses 1.0 mg/day vs 2.0 mg/day
dose.subset2 <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0, 2.0))
ttest2 <- t.test(len~dose, data = dose.subset2)
ttest2
```

The p-values for both tests are much smaller than 0.05, and all CIs do not overlap with 0, suggesting that Vitamin C dose has an impact on tooth growth. 

However, since we did multiple comparisons, let's do a Bonferroni correction to see if our results still hold. 
```{r}
# Store all p-values in a vector
pvals <- c(ttest1$p.value, ttest2$p.value)

# Do Bonferroni correction
adjusted.pvals <- p.adjust(pvals, method = "bonferroni")

# Calculate number of significant p-values
sum(adjusted.pvals < 0.05)
```

After Bonferroni correction, both p-values are still significant. So, our results still hold. 

## Conclusions

Assuming that (1) the ToothGrowth data sample is representative of the population, and (2) the distribution of the sample means is approximately normal, we can conclude the following: 
1. Delivery method of Vitamin C does not have an impact on tooth growth. 
2. Vitamin C dose does affect tooth growth. Specifically, the higher the dose, the larger the tooth growth.   
