---
author: "Rebecca Kee"
date: "8/10/2019"
output: 
    html_document: 
        keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Regression Models Quiz 3 

Q1. Consider the `r mtcars` data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as confounder. Give the adjusted estimate for the expected change in mpg comparing 8 cylinders to 4.
```{r}
data(mtcars)
fit <- lm(mpg ~ factor(cyl) + wt, data = mtcars)
summary(fit)$coefficients
```
The adjusted estimate is -6.071. 


Q2. Consider the `r mtcars` data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as a possible confounding variable. Compare the effect of 8 versus 4 cylinders on mpg for the adjusted and unadjusted by weight models. Here, adjusted means including the weight variable as a term in the regression model and unadjusted means the model without weight included. What can be said about the effect comparing 8 and 4 cylinders after looking at models with and without weight included?
```{r}
# Fit model without adjusting for weight and obtain coefficient comparing 8 and 4 cylinders
fit2 <- lm(mpg ~ factor(cyl), data = mtcars)
summary(fit2)$coefficient[3]
```
```{r}
# Retrieve coefficient comparing 8 an 4 cylinders from the first adjusted model
summary(fit)$coefficient[3]
```
Answer: Holding weight constant, cylinder appears to have less of an impact on mpg than if weight is disregarded.

Q3. Consider the `r mtcars` data set. Fit a model with mpg as the outcome that considers number of cylinders as a factor variable and weight as confounder. Now fit a second model with mpg as the outcome model that considers the interaction between number of cylinders (as a factor variable) and weight. Give the P-value for the likelihood ratio test comparing the two models and suggest a model using 0.05 as a type I error rate significance benchmark.
```{r}
# Fit adjusted model, no interation 
fit <- lm(mpg ~ factor(cyl) + wt, data = mtcars)
summary(fit)
```
```{r}
# Fit adjusted model, with interation 
fit3 <- lm(mpg ~ factor(cyl) * wt, data = mtcars)
summary(fit3)
```
```{r}
# Compare both models
anova(fit, fit3)
```
Answer: The P-value is larger than 0.05. So, according to our criterion, we would fail to reject, which suggests that the interaction terms may not be necessary.

Q4.Consider the `r mtcars` data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight inlcuded in the model as 
```{r}
fit4 <- lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars)
summary(fit4)
```
How is the wt coefficient interpreted?

Answer: The estimated expected change in MPG per one ton increase in weight for a specific number of cylinders (4, 6, 8).

Q5. Consider the following data set
```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
```

Give the hat diagonal for the most influential point
```{r}
fit5 <- lm(y ~ x)
hatvalues(fit5)
```
The hat value for the most influential point is 0.9946. 

Q6. Consider the following data set
```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)

# Note: This is the exact same data set used in Q5.
```

Give the slope dfbeta for the point with the highest hat value.
```{r}
dfbetas(fit5)
```
The dfbeta for the point with highest hat value is -134.

Q7. Consider a regression relationship between Y and X with and without adjustment for a third variable Z. Which of the following is true about comparing the regression coefficient between Y and X with and without adjustment for Z.

Answer: It is possible for the coefficient to reverse sign after adjustment. For example, it can be strongly significant and positive before adjustment and strongly significant and negative after adjustment.
